Intro to R - Module 2
========================================================

In this module, you'll learn how to create sequences of numbers in R. Then, you'll apply that knowledge to take subsets of vectors.

---

The simplest way to create a sequence of numbers in R is by using the `:` operator. Type `1:20` to see how it works.

```{r}
1:20
```

---

That gave us every integer between (and including) 1 and 20. We could also use it to create a sequence of real numbers. For example, try `pi:10`.

```{r}
pi:10
```

---

The result is a vector of real numbers starting with pi (3.142...) and increasing in increments of 1. The upper limit of 10 is never reached, since the next number in our sequence would be greater than 10.

---

What happens if we do this: `15:1`? Give it a try to find out.

```{r}
15:1
```

---

It counted backwards in increments of 1! It's unlikely we'd want this behavior, but nonetheless it's good to know how it could happen.

---

Generally, we'll desire more control over a sequence we're creating. The `seq()` function serves this purpose.

---

The most basic use of `seq()` does exactly the same thing as the `:` operator. Try `seq(1, 20)` to see this.

```{r}
seq(1, 20)
```

This gives us the same output as `1:20`. However, let's say that instead we want a vector of numbers ranging from 0 to 10, incremented by 0.5. `seq(0, 10, by=0.5)` does just that. Try it out.

```{r}
seq(0, 10, by=0.5)
```

---

Or maybe we don't care what the increment is and we just want a sequence of 30 numbers between 5 and 10. `seq(5, 10, length=30)` does the trick. Give it shot now and store the result in a new variable called `my_seq`.

```{r}
my_seq <- seq(5, 10, length=30)
```

---

To confirm that `my_seq` has length 30, we can use the `length()` function. Try it now.

```{r}
length(my_seq)
```

---

Let's pretend we don't know the length of `my_seq`, but we want to generate a sequence of integers from 1 to N, where N represents the length of the `my_seq` vector. In other words, we want a new vector (1, 2, 3, ...) that is the same length as `my_seq`.

---

There are several ways we could do this. One possibility is to combine the `:` operator and the `length()` function like this: `1:length(my_seq)`. Give that a try.

```{r}
1:length(my_seq)
```

---

However, as is the case with many common tasks, R has a built-in function for this purpose called `seq_along()`. Type `seq_along(my_seq)` to see it in action.

```{r}
seq_along(my_seq)
```

---

There are often several approaches to solving the same problem, particularly in R. Simple approaches that involve less typing are generally best. It's also important for your code to be readable, so that you and others can figure out what's going on without too much trouble.

---

If R has a built-in function for a particular task, it's likely that function is highly optimized for that purpose and is your best option. As you become a more advanced R programmer, you'll want to design your own functions to perform tasks when there are no better options. We'll explore writing your own functions in future lessons.

---

